---
title: "LabBook_21_10_2016"
author: "Claire Green"
output: pdf_document
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))
```

#Tuesday

I need to do some catching up as a lot of September/October has been theoretical and I haven't got to grips with implementing the investigations I planned way back in August. In the previous analysis I did an experiment to look at the similarities in the expression patterns of my DEGs in the different datasets. Although I had hoped mutation backgrounds would cluster, instead there appears to be a platform effect that is stronger. To confirm this finding, I conducted an experiment to include the controls - the aim to be that controls would cluster together and disease separately. The results were not entirely clear, but it does appear there is still a platform effect. 

This is the R script I used:

```{r, eval = FALSE}
#Selecting DEGS from expression matrix

#Load list of interesting genes
#setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/GeneExpressionAnalysis/Microarray/")
setwd(dir = "/Users/clairegreen/Desktop/")
Genelist <- read.csv("overlap_ens2hgnc_4RNAseq.csv", header = TRUE)

#load dataset
setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/DEG_Test2/")
exprs <- read.csv("C9rankeduniqueresult.csv")

#Make gene symbol row names
#rownames(exprs) <- exprs$Ensembl
exprspat <- exprs[,49:51]
exprspat[,(length(exprspat)+1)] <- exprs$Ensembl

#Make gene symbol a column
# exprspat <- cbind(exprspat, exprs$Gene.Symbol)
# colnames(exprspat)[length(exprspat)] <- "Gene.Symbol"

#Merge by interesting gene names with expression to form matrix
patgene <- merge(Genelist, exprspat, by.x = "ensembl_gene_id", by.y = "V4")
#patgene <- patgene[!duplicated(patgene[,11]),]
# rownames(patgene) <- patgene$V1
# patgene[,1] <- NULL

setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/Pathways_to_TDP-43/Co-expression/")
write.csv(patgene, file = "VCP_DEG_CON_ens.csv")

#### Cor.test Method ####

library(tictoc)
library(gdata)

##load dataset
setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/Pathways_to_TDP-43/Co-expression/")
Exprs_val <- read.csv("C9r_DEG_CON_Exprs.csv")
rownames(Exprs_val) <- Exprs_val[,2]
Exprs_val[,1:2] <- NULL

CorExprMat <- t(Exprs_val)

reg <- matrix(0, ncol(CorExprMat), ncol(CorExprMat))

tic()
for (i in 1:ncol(CorExprMat)){
  for (j in 1:ncol(CorExprMat)){
    reg[i,j] <- cor.test(CorExprMat[,i], CorExprMat[,j], method = "spearman")$estimate
  }}

rownames(reg) <- colnames(reg) <- colnames(CorExprMat)
toc()


#Extract R values
corRadj <- reg
corRadj[lower.tri(corRadj, diag = TRUE)] <- NA

#Turn into vector
corRadj <- as.matrix(corRadj)
corRvec <- unmatrix(corRadj)
#Remove NA values
corRvec <- na.omit(corRvec)
corRvec <- as.data.frame(corRvec)

write.csv(corRvec, file = "C9r_CON_CorResults.csv")


#Generate matrix with all Rho values

C9mR <- read.csv("C9m_CorResults.csv")
CHmR <- read.csv("CHMP2B_CorResults.csv")
GRNR <- read.csv("GRN_CorResults.csv")
VCPR <- read.csv("VCP_CorResults.csv")
C9rR <- read.csv("C9r_CorResults.csv")

C9mCON <- read.csv("C9m_CON_CorResults.csv")
C9rCON <- read.csv("C9r_CON_CorResults.csv")
CHCON <- read.csv("CH_CON_CorResults.csv")
GRNCON <- read.csv("FTLD_CON_CorResults.csv")
VCPCON <- read.csv("VCP_CON_CorResults.csv")



RhoValues <- merge(C9mR, CHmR, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, GRNR, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, VCPR, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, C9rR, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, C9mCON, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, C9rCON, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, CHCON, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, GRNCON, by.x = "X", by.y = "X")
RhoValues <- merge(RhoValues, VCPCON, by.x = "X", by.y = "X")

colnames(RhoValues) <- c("GenePair","C9m", "CHMP2B", "GRN", "VCP", "C9r", "C9mCON", "C9rCON",
                         "CHCON", "GRNCON", "VCPCON")

rownames(RhoValues) <- RhoValues$GenePair
RhoValues[,1] <- NULL

Rho <- matrix(0, ncol(RhoValues), ncol(RhoValues))

tic()
for (i in 1:ncol(RhoValues)){
  for (j in 1:ncol(RhoValues)){
    Rho[i,j] <- cor.test(RhoValues[,i], RhoValues[,j], method = "kendall")$p.value
  }}

rownames(Rho) <- colnames(Rho) <- colnames(RhoValues)
toc()


#Extract R values
corRadj <- Rho
corRadj[lower.tri(corRadj, diag = TRUE)] <- NA

#Turn into vector
corRadj <- as.matrix(corRadj)
corRvec <- unmatrix(corRadj)
#Remove NA values
corRvec <- na.omit(corRvec)
corRvec <- as.data.frame(corRvec)

write.csv(corRvec, file = "Alldatasets_Pval.csv")

```

Results table can be found in the file "Alldatasets_Correlation.csv":

         
| Dataset1 | Dataset2 | corRvec      | pval        | Relationship |
|----------|----------|--------------|-------------|--------------|
| GRN      | GRNCON   | 0.230625858  | 0           | Correlation  |
| C9r      | C9rCON   | 0.194266757  | 7.36E-277   | Correlation  |
| CHCON    | GRNCON   | 0.185907259  | 3.53E-256   | Correlation  |
| GRN      | CHCON    | 0.168137597  | 1.97E-206   | Correlation  |
| C9m      | CHCON    | 0.164071069  | 3.45E-202   | Correlation  |
| C9rCON   | GRNCON   | 0.115052331  | 7.22E-100   | Correlation  |
| C9r      | CHCON    | 0.110565899  | 1.38E-90    | Correlation  |
| C9m      | C9mCON   | 0.116612111  | 1.54E-85    | Correlation  |
| C9mCON   | CHCON    | 0.116942577  | 6.05E-85    | Correlation  |
| C9r      | GRNCON   | 0.106495231  | 2.53E-84    | Correlation  |
| CHMP2B   | CHCON    | 0.113398182  | 1.41E-78    | Correlation  |
| C9m      | GRNCON   | 0.101396028  | 1.41E-78    | Correlation  |
| VCP      | C9rCON   | 0.099139702  | 4.33E-74    | Correlation  |
| C9mCON   | GRNCON   | 0.104771445  | 1.13E-68    | Correlation  |
| GRN      | C9rCON   | 0.093170481  | 4.65E-65    | Correlation  |
| CHMP2B   | C9r      | 0.078650136  | 3.00E-38    | Correlation  |
| C9m      | GRN      | 0.070308169  | 4.34E-38    | Correlation  |
| C9rCON   | CHCON    | 0.069934314  | 5.62E-38    | Correlation  |
| C9m      | C9r      | 0.064566617  | 1.87E-32    | Correlation  |
| VCP      | GRNCON   | 0.061676982  | 1.18E-29    | Correlation  |
| CHMP2B   | GRNCON   | 0.06808106   | 1.70E-29    | Correlation  |
| CHMP2B   | C9mCON   | 0.07348021   | 2.12E-28    | Correlation  |
| C9m      | CHMP2B   | 0.064481371  | 6.62E-27    | Correlation  |
| GRN      | C9r      | 0.058314719  | 4.36E-26    | Correlation  |
| GRN      | VCP      | 0.057055619  | 3.27E-25    | Correlation  |
| C9mCON   | VCPCON   | 0.066445802  | 3.11E-23    | Correlation  |
| C9r      | C9mCON   | 0.059344868  | 7.09E-23    | Correlation  |
| C9r      | VCPCON   | 0.041635816  | 1.04E-11    | Correlation  |
| CHCON    | VCPCON   | 0.040571495  | 2.56E-11    | Correlation  |
| GRN      | C9mCON   | 0.036583101  | 1.33E-09    | Correlation  |
| CHMP2B   | C9rCON   | 0.034647364  | 8.99E-09    | Correlation  |
| VCP      | C9mCON   | 0.032794016  | 4.74E-08    | Correlation  |
| VCP      | C9r      | 0.02915095   | 1.12E-07    | Correlation  |
| VCP      | VCPCON   | 0.021827297  | 0.000346285 | Correlation  |
| GRN      | VCPCON   | 0.021566726  | 0.000433578 | Correlation  |
| VCP      | CHCON    | 0.01886406   | 0.00054837  | Correlation  |
| C9m      | VCPCON   | 0.018218401  | 0.002575407 | Correlation  |
| GRNCON   | VCPCON   | 0.014362344  | 0.018112961 | Correlation  |
| C9m      | C9rCON   | 0.011292094  | 0.036304906 | Correlation  |
| C9rCON   | VCPCON   | -0.009142945 | 0.131812358 | Correlation  |
| C9mCON   | C9rCON   | 0.007806611  | 0.191165437 | Correlation  |
| CHMP2B   | VCPCON   | 0.005023414  | 0.456970162 | Correlation  |
| CHMP2B   | VCP      | 0.002826308  | 0.64094965  | Correlation  |
| CHMP2B   | GRN      | -0.000809281 | 0.894258976 | Correlation  |
| C9m      | VCP      | 0.000608703  | 0.910640153 | Correlation  |


![](/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/Pathways_to_TDP-43/Co-expression/DatasetCorrelation/Cytoscape_datasetcorrelation.pdf)

Cytoscape output shows that it seems like platform is having too much of an effect. 