---
title: "LabBook_09_09_2016"
author: "Claire Green"
output: pdf_document
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))
```

#Thursday

I'm beginning to conduct the first part of my analysis. This involves generating coexpression values for the commonly DEGs for each dataset. For full explanation, refer to the document "Pathways to TDP"

First, I needed to extract the DEG expression values for each dataset. That requires running the following script, adjusted for each dataset:

```{r, eval=FALSE}

#Selecting DEGS from expression matrix

#Load list of interesting genes
setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/GeneExpressionAnalysis/Microarray/")
Genelist <- read.csv("MADEGs.csv")

#load dataset
setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/GeneExpressionAnalysis/Microarray/TopGenes_2016-02-15")
exprs <- read.csv("C9rankeduniqueresult.csv")

#Make gene symbol row names
rownames(exprs) <- exprs$Gene.Symbol
exprspat <- exprs[,52:59]

#Make gene symbol a column
exprspat <- cbind(exprspat, exprs$Gene.Symbol)
colnames(exprspat)[length(exprspat)] <- "Gene.Symbol"

#Merge by interesting gene names with expression to form matrix
patgene <- merge(Genelist, exprspat, by.x = "Gene", by.y = "Gene.Symbol")
rownames(patgene) <- patgene$Gene
patgene[,1] <- NULL

setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/Pathways_to_TDP-43/Co-expression/")
write.csv(patgene, file = "C9_DEG_Exprs.csv")
```

This produced a .csv file for each of the 5 microarray datasets containing the expression values for each of the DEGs for only patients. 

For the PET dataset I had to adjust the code as the gene names were different. I used the list generated from the consensus analysis and ran the code with this instead.

```{r, eval=FALSE}
#Selecting DEGS from expression matrix

#Load list of interesting genes
setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/GeneExpressionAnalysis/Microarray/")
setwd(dir = "/Users/clairegreen/Desktop/")
Genelist <- read.table("My TDP-43 Genes.txt")

#load dataset
setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/DEG_Test2/")
exprs <- read.csv("PETNOrankeduniqueresult.csv")

#Make gene symbol row names
rownames(exprs) <- exprs$X
exprspat <- exprs[,9:15]

#Make gene symbol a column
exprspat <- cbind(exprspat, exprs$X)
colnames(exprspat)[length(exprspat)] <- "Gene.Symbol"

#Merge by interesting gene names with expression to form matrix
patgene <- merge(Genelist, exprspat, by.x = "V1", by.y = "Gene.Symbol")
rownames(patgene) <- patgene$V1
patgene[,1] <- NULL

setwd(dir = "/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/Pathways_to_TDP-43/Co-expression/")
write.csv(patgene, file = "C9_DEG_Exprs.csv")
```


Next, I needed to run correlation analysis to find coexpression values for each gene pair

Example code

```{r, eval=FALSE}
#### Cor.test Method ####

library(tictoc)
library(gdata)

##load dataset
Exprs_val <- read.csv("CHMP2B_DEG_Exprs.csv")
rownames(Exprs_val) <- Exprs_val[,1]
Exprs_val[,1] <- NULL

CorExprMat <- t(Exprs_val)

reg <- matrix(0, ncol(CorExprMat), ncol(CorExprMat))

tic()
for (i in 1:ncol(CorExprMat)){
  for (j in 1:ncol(CorExprMat)){
    reg[i,j] <- cor.test(CorExprMat[,i], CorExprMat[,j], method = "spearman")$estimate
  }}

rownames(reg) <- colnames(reg) <- colnames(CorExprMat)
toc()


#Extract R values
corRadj <- reg
corRadj[lower.tri(corRadj, diag = TRUE)] <- NA

#Turn into vector
corRadj <- as.matrix(corRadj)
corRvec <- unmatrix(corRadj)
#Remove NA values
corRvec <- na.omit(corRvec)
corRvec <- as.data.frame(corRvec)

write.csv(corRvec, file = "CHMP2B_CorResults.csv")

```

