---
title: "LabBook_13_03_2017"
author: "Claire Green"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Monday
I need to work out STILL a way of assigning the threshold to use for my DEG overlap analysis. This has been plaguing me for a long time and I really don't know exactly how I'm going to tackle it. The spike in method that I reported previously works really well, the problem is the fact I can't enrich for disease genes (because they are in the benchmarking list) but also that I CAN'T GET THE SAME RESULTS AGAIN. I even went to the trouble of copying and pasting the script as it was to make sure I used the exact same one, but for some reason something has changed and now it's saying for the same analysis that the optimum threshold is in the 7000s

Here are the methods I have tried so far with the positives and negatives:

1. Assigned threshold   
This was my first method - just take a top threshold of N genes and see what overlaps. This gave relatively good results but the problem was that I knew there was no evidence behind the threshold I had chosen. I thought - there must be a threshold at which my enrichment signal is clearest. Which led me to method number two...

2. Enrichment spike in   
This involved making a meta-list of my benchmarks. This was then used to run enrichment analysis at different thresholds so as to define the exact threshold at which enrichment was highest. This was at first 5996. The positives were that the enrichment of the processes on EnrichR were vastly improved, as were, obviously, the benchmark enrichments. However there were two big problems: 1, I couldn't repeat the 5996 results (even runninhg exaclty the same script gave me a result of 7000 or so) and 2, If I used the benchmarks as the spike-in list then I couln't use it as a bnechmark list - it would be a circular argument. I attempted to cut down the spike-in list to just TDP-43 PPI genes but the problem then was that any functional enrichment analysis I did would be biased towards TDP-43 interactors which enrich heavily for mRNA processing. Therefore any mRNA processing enrichment is skewed towards the thresholding. I did try the enrichment without the overlapping genes which did still (sort of) enrich for mRNA processing but it was far weaker. This was a good back up if the following didn't work.

3. Heatmap   
Dennis suggested that a way of identifying up and downregulated genes could be to use the dendrogram of a heatmap to cluster the similarly expressed genes together. 

```{r}
setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/noMedian/")

C9 <- read.csv("C9_unique.csv")
CH <- read.csv("CH_unique.csv")
sals <- read.csv("sals_unique.csv")
ftld <- read.csv("ftld_unique.csv")
vcp <- read.csv("vcp_unique.csv")

setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/TDP-43_DEseq2/")

pet <- read.csv("PET_results_keepfiltering.csv")
pet <- pet[!duplicated(pet$hgnc_symbol),]
rav <- read.csv("RAV_results_keepfiltering.csv")
rav <- rav[!duplicated(rav$hgnc_symbol),]

#Find genes that are present across all FULL datasets
C9gen <- C9$Gene.Symbol
CHgen <- CH$Gene.Symbol
salsgen <- sals$Gene.Symbol
ftldgen <- ftld$Gene.Symbol
vcpgen <- vcp$Gene.Symbol
petgen <- pet$hgnc_symbol
ravgen <- rav$hgnc_symbol

genelist <- Reduce(intersect, list(C9gen, CHgen, salsgen, ftldgen, vcpgen, petgen, ravgen))

#take that subset from each of the datasets
subsetC9 <- subset(C9, C9$Gene.Symbol %in% genelist, drop = TRUE)
rownames(subsetC9) <- subsetC9$Gene.Symbol
subsetC9[,1] = NULL
subsetC9 <- subsetC9[order(row.names(subsetC9)),]

subsetCH <- subset(CH, CH$Gene.Symbol %in% genelist, drop = TRUE)
rownames(subsetCH) <- subsetCH$Gene.Symbol
subsetCH[,1] = NULL
subsetCH <- subsetCH[order(row.names(subsetCH)),]

subsetsals <- subset(sals, sals$Gene.Symbol %in% genelist, drop = TRUE)
rownames(subsetsals) <- subsetsals$Gene.Symbol
subsetsals[,1] = NULL
subsetsals <- subsetsals[order(row.names(subsetsals)),]

subsetftld <- subset(ftld, ftld$Gene.Symbol %in% genelist, drop = TRUE)
rownames(subsetftld) <- subsetftld$Gene.Symbol
subsetftld[,1] = NULL
subsetftld <- subsetftld[order(row.names(subsetftld)),]

subsetvcp <- subset(vcp, vcp$Gene.Symbol %in% genelist, drop = TRUE)
rownames(subsetvcp) <- subsetvcp$Gene.Symbol
subsetvcp[,1] = NULL
subsetvcp <- subsetvcp[order(row.names(subsetvcp)),]

subsetpet <- subset(pet, pet$hgnc_symbol %in% genelist, drop = TRUE)
rownames(subsetpet) <- subsetpet$Gene.Symbol
subsetpet[,1] = NULL
subsetpet <- subsetpet[order(subsetpet$hgnc_symbol),]

subsetrav <- subset(rav, rav$hgnc_symbol %in% genelist, drop = TRUE)
rownames(subsetrav) <- subsetrav$Gene.Symbol
subsetrav[,1] = NULL
subsetrav <- subsetrav[order(subsetrav$hgnc_symbol),]


#Generate a matrix with gene names and log fold change values from each
LFC <- data.frame(gene=row.names(subsetC9), 
                C9orf72=subsetC9$logFC,
                CHMP2B=subsetCH$logFC,
                sALS.1=subsetsals$logFC, 
                FTLD=subsetftld$logFC,
                VCP=subsetvcp$logFC, 
                C9sALS=subsetpet$log2FoldChange, 
                sALS.2=subsetrav$log2FoldChange)
rownames(LFC) <- LFC$gene
LFC[,1] <- NULL
LFC <- as.matrix(LFC)

LFCNC <- data.frame(gene=row.names(subsetC9), 
                   C9orf72=subsetC9$logFC,
                   sALS=subsetsals$logFC, 
                   FTLD=subsetftld$logFC,
                   VCP=subsetvcp$logFC, 
                   C9_sALS=subsetpet$log2FoldChange, 
                   sALS=subsetrav$log2FoldChange) 

rownames(LFCNC) <- LFCNC$gene
LFCNC[,1] <- NULL
LFCNC <- as.matrix(LFCNC)

library(gplots)
library(heatmap3)
hmcols<- colorRampPalette(c("green4","green","white", "red","red4"))(256)
# 
# Platform <- c("orange","orange","orange","orange","blue", "blue")
# Tissue <- c("magenta", "magenta", "yellow", "cyan", "yellow", "magenta")

colsidecolorsLFCNC <- cbind(Platform=c("darkgoldenrod3","darkgoldenrod3","darkgoldenrod3","darkgoldenrod3","darkgoldenrod1", "darkgoldenrod1"), 
                       Disease=c("dodgerblue4","dodgerblue4","cyan","cornflowerblue","dodgerblue4","dodgerblue4"),
                       Tissue=c("magenta", "magenta", "purple", "purple4", "purple", "magenta"))

colsidecolorsLFC <- cbind(Disease=c("dodgerblue4","dodgerblue4","dodgerblue4","lightblue","cornflowerblue","dodgerblue4","dodgerblue4"),
                          Platform=c("darkgoldenrod3","darkgoldenrod3","darkgoldenrod3","darkgoldenrod3","darkgoldenrod3","darkgoldenrod1", "darkgoldenrod1"),
                          Tissue=c("magenta","magenta", "magenta", "purple", "purple4", "purple", "magenta"))

heatmap3(LFC, 
         col = hmcols,
         Rowv = TRUE,
         Colv = TRUE,
         distfun = dist,
         hclustfun = hclust,
         scale = "row",
         labCol = colnames(LFC),
         ColSideColors = colsidecolorsLFC,
         ColSideWidth = 1,
         cexCol = 1.2,
         cexRow = 0.01,
         xlab = "Datasets",
         ylab = "Genes")

heatmap3(LFCNC, 
         col = hmcols,
         Rowv = TRUE,
         Colv = TRUE,
         distfun = dist,
         hclustfun = hclust,
         scale = "row",
         labCol = colnames(LFCNC),
         ColSideColors = colsidecolorsLFCNC,
         ColSideWidth = 1,
         cexCol = 1.2,
         cexRow = 0.01,
         xlab = "Datasets",
         ylab = "Genes")
```
   
   
   As you can see from the first heatmap, the CHMP2B dataset has really odd expression in comparison to the others. The colours are very bright and almost opposite to every other sample. At this point although I decided the heatmap wouldn't necessarily be the best option, I don't think including the CHMP2B dataset any further is a good idea. This will change a lot of results...
\newpage

3. Fold change   
My thought then was to try and use a measure that was completely unrelated to the benchamrk list. I wanted to avoid p value as it's so scrutinised and arbitrary so I thought I would try fold change instead. On Friday I began these tets but the problem was that for log2FC I needed such a low threshold that I thought it looked a bit dodgy. I'm not really familiar with the field but using a threshold of 0.1 felt not so great. I don't know if this really makes a difference but I decided to use fold change adjusted to a scale centered around 0.

```{r, eval=FALSE}
setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/noMedian/")

C9 <- read.csv("C9_unique.csv")
C9 <- C9[order(C9$P.Value),]
CH <- read.csv("CH_unique.csv")
CH <- CH[order(CH$P.Value),]
sals <- read.csv("sals_unique.csv")
sals <- sals[order(sals$P.Value),]
ftld <- read.csv("ftld_unique.csv")
ftld <- ftld[order(ftld$P.Value),]
vcp <- read.csv("vcp_unique.csv")
vcp <- vcp[order(vcp$P.Value),]

setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/TDP-43_DEseq2/")

pet <- read.csv("PET_results_keepfiltering.csv")
pet <- pet[!duplicated(pet$hgnc_symbol),]
pet$"FoldChange"<-2^pet$log2FoldChange
pet$"FoldChange"[pet$"FoldChange"<1]<-(-1)/pet$"FoldChange"[pet$"FoldChange"<1]
rav <- read.csv("RAV_results_keepfiltering.csv")
rav <- rav[!duplicated(rav$hgnc_symbol),]
rav$"FoldChange"<-2^rav$log2FoldChange
rav$"FoldChange"[rav$"FoldChange"<1]<-(-1)/rav$"FoldChange"[rav$"FoldChange"<1]

thresh <- 1

upC9 <- subset(C9, C9$Fold.Change >= thresh)
upC9gene <- upC9$Gene.Symbol

# upCH <- subset(CH, CH$Fold.Change >= thresh)
# upCHgene <- upCH$Gene.Symbol

upSALS <- subset(sals, sals$Fold.Change >= thresh)
upSALSgene <- upSALS$Gene.Symbol

upFTLD <- subset(ftld, ftld$Fold.Change >= thresh)
upFTLDgene <- upFTLD$Gene.Symbol

upVCP <- subset(vcp, vcp$Fold.Change >= thresh)
upVCPgene <- upVCP$Gene.Symbol

upPET <- subset(pet, pet$FoldChange >= thresh)
upPETgene <- upPET$hgnc_symbol

upRAV <- subset(rav, rav$FoldChange >= thresh)
upRAVgene <- upRAV$hgnc_symbol

INTUP <- Reduce(intersect, list(upC9gene, upSALSgene, upFTLDgene, upVCPgene, upPETgene, upRAVgene))

cat(INTUP, sep = "\n")

setwd("/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/FoldChangeResults")
write.table(INTUP, "intersect_up_1.txt", col.names = F, row.names = F, quote = F)



#### DOWN ####
thresh <- -1

downC9 <- subset(C9, C9$Fold.Change <= thresh)
downC9gene <- downC9$Gene.Symbol

# downCH <- subset(CH, CH$Fold.Change <= thresh)
# downCHgene <- downCH$Gene.Symbol

downSALS <- subset(sals, sals$Fold.Change <= thresh)
downSALSgene <- downSALS$Gene.Symbol

downFTLD <- subset(ftld, ftld$Fold.Change <= thresh)
downFTLDgene <- downFTLD$Gene.Symbol

downVCP <- subset(vcp, vcp$Fold.Change <= thresh)
downVCPgene <- downVCP$Gene.Symbol

downPET <- subset(pet, pet$FoldChange <= thresh)
downPETgene <- downPET$hgnc_symbol

downRAV <- subset(rav, rav$FoldChange <= thresh)
downRAVgene <- downRAV$hgnc_symbol

INTDOWN <- Reduce(intersect, list(downC9gene, downSALSgene, downFTLDgene, downVCPgene, downPETgene, downRAVgene))

setwd("/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/FoldChangeResults")
write.table(INTDOWN, "intersect_down_1.txt", col.names = F, row.names = F, quote = F)

cat(INTDOWN, sep = "\n")
```

As you can see, I decided to remove the CHMP2B dataset. Taking upregulated genes with FC above or equal to 1, and downregulated genes as below or equal to -1, I ended up with 329 upregulated genes and 69 downregulated genes.

\newpage

Upregulated Genes

| 1-30    | 31-60   | 61-90   | 91-120   | 121-15  | 151-180 | 181-210 | 211-240   | 241-270  | 271-300 | 301-329 |
|---------|---------|---------|----------|---------|---------|---------|-----------|----------|---------|---------|
| RALGDS  | CTBP2   | PIKFYVE | ZNF573   | ABCA8   | TTF1    | AGO3    | SETD2     | NBR1     | SMC2    | CHPT1   |
| AHCYL1  | CALD1   | GPNMB   | HIPK1    | TMED10  | WDR11   | ALMS1   | CREG1     | UNC13B   | ENPP2   | METTL9  |
| EEF1A1  | PAPOLA  | PABPC1  | ZNF12    | HSBP1   | MYOF    | RANGRF  | CEP57     | FAM65B   | PHKB    | BCLAF1  |
| RCN1    | NOTCH2  | SMC5    | RND3     | PPFIA1  | SRRM1   | YY1AP1  | RPRD1A    | MARK3    | ZNF141  | POT1    |
| STOM    | SLCO3A1 | VCAN    | PRPF40A  | AASS    | CLK4    | GIN1    | ZFX       | MAP4K5   | XBP1    | ZNF226  |
| IFITM2  | LPL     | PHLDA1  | SNAPC3   | TRIT1   | ADAMTS9 | ABL1    | HNRNPA2B1 | PPP1R12B | MAP3K7  | RBM12   |
| KCTD20  | CCNL1   | CCNL2   | ITPR2    | GNA13   | KDM5A   | PTK2    | LIMA1     | SH3GLB1  | ERP44   | TGOLN2  |
| CTSH    | SMAD5   | AMD1    | LPP      | SPEN    | THUMPD2 | PIGA    | MON2      | PLEKHA2  | RSAD1   | LRCH3   |
| MAFF    | ATG14   | GSPT1   | CDH11    | CMTM6   | PIK3C2A | LBR     | ZCCHC8    | EIF3E    | DHX29   | C18orf8 |
| LYN     | ZFYVE26 | EPS8    | ZCCHC6   | SRSF4   | ANKRD10 | SLK     | UPF3B     | UBXN7    | NAA50   | CENPU   |
| PRPF4B  | PPP4R2  | N4BP2L2 | PNN      | TARDBP  | PRR11   | SUCO    | TIMM8A    | CLIC4    | AGO4    | ATP8B1  |
| CNOT6   | SASH1   | TBC1D2B | AUTS2    | KATNBL1 | ABI1    | MARCH3  | SORBS1    | ZCCHC10  | AKR1C1  | ZNF611  |
| SGK1    | ZBTB10  | NSMAF   | PUM2     | HCFC2   | RSL24D1 | STK3    | GSAP      | SLC26A2  | TMEM165 | TMEM38B |
| EDNRB   | MLH3    | KCTD12  | CRISPLD2 | ICE1    | TMX1    | SF3A1   | NBPF1     | USP3     | AHR     | RFX3    |
| KAT6A   | NUDT21  | TMCO1   | TIMP2    | DERL1   | ADAMTS1 | ANKRD28 | TENM1     | C6orf120 | STX2    | SAV1    |
| SPARC   | DDX21   | NKTR    | ESF1     | PCSK5   | USP1    | TNFAIP3 | KRIT1     | KIAA0141 | DIDO1   | KLF3    |
| RBM6    | HNRNPA3 | NR2C1   | RPS6KA2  | NUP43   | ERBB4   | PXDN    | SREK1     | MED14    | FLNC    | WRN     |
| ASAH1   | PSAP    | ROCK1   | YTHDF1   | PRPSAP1 | PTPRA   | JMJD1C  | WTAP      | CPOX     | PXMP4   | HNMT    |
| RBPJ    | ALOX5   | RNASEH1 | GOLGA7   | SCAF11  | MAP4K3  | CRLF3   | NEK1      | ANXA1    | AKAP13  | TMEM5   |
| ATF7IP  | BTG1    | CBFB    | SEC24A   | GNAI3   | SLC11A2 | ZFHX3   | REL       | WSB1     | PRDX4   | SSR2    |
| CRY1    | IFI16   | MCMBP   | PALLD    | CHSY1   | RAP1B   | CDK13   | STRN3     | FHL1     | RCN2    | SRSF10  |
| PHIP    | CREB3L2 | MFN1    | PKIA     | RAB27A  | IFRD1   | ZNF83   | SLC33A1   | SH2B3    | CBLB    | TRIM5   |
| FAM129A | TRAK1   | ZNF146  | SMAD1    | ARMC1   | SSFA2   | PHF21A  | SOCS2     | HEXB     | SLC35A3 | NCOA6   |
| DDX18   | KCNE4   | LRBA    | CPSF7    | PDE4B   | HMGN4   | PROSER1 | ZNF536    | TCF12    | TWIST1  | RFX7    |
| PLOD2   | PHF3    | COL4A1  | ZBTB5    | CARD8   | PDS5A   | PTPRC   | POGZ      | TPP1     | DENND4A | STAG2   |
| SERBP1  | PTP4A3  | ZNF292  | RAP1A    | HLA-B   | KLF6    | MAN2A1  | AKR1B1    | TRIM38   | CD44    | CPSF6   |
| TMEM43  | HCLS1   | SOAT1   | STX16    | GCH1    | RBBP6   | ITGA6   | DIP2A     | OGT      | FOXN3   | REV3L   |
| SEPT8   | PCNA    | LAPTM5  | EFEMP1   | ZFC3H1  | RBM5    | ADNP    | PLSCR1    | MUS81    | SENP7   | CCDC90B |
| MED13L  | CNOT1   | AMPD3   | SAT1     | ALPK1   | SUN1    | TWSG1   | U2SURP    | CHD1     | ATM     | SMNDC1  |
| ADAM17  | NEK4    | PNISR   | CEBPZ    | CDK12   | ANAPC5  | HEATR3  | ASXL2     | RIT1     | BRAP    |         |

\newpage

Downregulated genes

| 1-30     | 31-60   | 61-69     |
|----------|---------|-----------|
| COX5A    | PEF1    | MARK4     |
| DRG1     | HABP4   | MED8      |
| SLC41A3  | WIPI2   | CXorf56   |
| SMARCA2  | KCNMB4  | DEDD      |
| VDAC1    | SAMM50  | CLCN6     |
| PHYH     | MGST3   | PSMF1     |
| PCDH17   | TPGS2   | LINC00094 |
| BNIP3    | CDH22   | TCEAL1    |
| RIMS1    | SPOCK1  | CYP2E1    |
| APLP2    | TSPYL5  |           |
| VDAC3    | GDI1    |           |
| PPP1R11  | FXR2    |           |
| SLC12A5  | SEPT6   |           |
| RPH3A    | LRRC14  |           |
| IMPDH1   | ALDOA   |           |
| GNB5     | UCHL1   |           |
| WASF1    | EIF1B   |           |
| PSMB2    | SIGMAR1 |           |
| ESRRG    | PMS2P1  |           |
| TMX4     | SSBP3   |           |
| ZBTB7A   | TMOD2   |           |
| PLCB1    | DOLPP1  |           |
| PPP3CA   | CEP170B |           |
| C16orf62 | R3HCC1  |           |
| KCNAB1   | SNX11   |           |
| AUH      | GIPC1   |           |
| OAT      | PDE4A   |           |
| CDIPT    | ANKRD27 |           |
| NDEL1    | TXN2    |           |
| CLCN3    | LPCAT4  |           |


